---
title: Dynamic HTML graph of Daily Temperatures
subtitle: Using DyGraph library.
week: 12
type: Case Study
reading:
   - Browse the [HTML Widgets page](http://gallery.htmlwidgets.org/) for many more examples. Take notes in your readme.md about potential uses in your project.
   - Explore the [DyGraphs webpage](http://rstudio.github.io/dygraphs/)
tasks:
   - Download daily weather data for Buffalo, NY using an API
   - Generate a dynamic html visualization of the timeseries. 
   - Save the graph to your project folder using Export->Save as Webpage
---


```{r setup, include=FALSE, purl=F}
source("functions.R")
source("knitr_header.R")
```

# Reading

```{r reading,results='asis',echo=F,purl=F}
md_bullet(rmarkdown::metadata$reading)
```


# Tasks

```{r tasks,results='asis',echo=F, purl=F}
md_bullet(rmarkdown::metadata$tasks)
```

## Background
In this session you will explore several ways to generate dynamic and interactive data displays.  These include making maps and graphs that you can pan/zoom, select features for more information, and interact with in other ways.  The most common output format is HTML, which can easily be embedded in a website (such as your final project!).

```{r cache=F, message=F,warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggmap)
library(htmlwidgets)
library(widgetframe)
```

If you don't have the packages above, install them in the package manager or by running `install.packages("widgetframe")`, etc. 

# Objective
> Make a dygraph of recent daily maximum temperature data from Buffalo, NY.

## Detailed Steps

First use the following code to download the daily weather data.

```{r, messages=F, warning=F, results="hide"}
library(tidyverse)
library(rnoaa)
library(xts)
library(dygraphs)
 
d=meteo_tidy_ghcnd("USW00014733",
                   date_min = "2016-01-01", 
                   var = c("TMAX"),
                   keep_flags=T) %>% 
   mutate(date=as.Date(date),
          tmax=as.numeric(tmax)/10) #Divide the tmax data by 10 to convert to degrees.
```

Remaining steps:

1. Convert `d` into an `xts` time series object using `xts()`.  You will need to specifify which column has the data (`d$tmax`) and `order.by=d$date`.   See `?xts` for help. 
2. Use `dygraph()` to draw the plot
3. Set the title of the dygraph to be `main="Daily Maximum Temperature in Buffalo, NY"`
4. Add a `dyRangeSelector()` with a `dateWindow` of `c("2020-01-01", "2020-10-31")`
5. Explore other options.  You could download another variable (precipitation?) and add it to the plot. Or imagine another way to visualize the data using one of the other interactive libraries.  

## Output

At a minimum, your final graph should look something like this:

```{r, echo=F, purl=F, warning=F, message=F}
library(widgetframe)

# Convert to a xts time series object as required by dygraph
dt=xts(d$tmax,order.by=d$date)

dygraph(dt, main = "Daily Maximum Temperature in Buffalo, NY") %>% 
  dyRangeSelector(dateWindow = c("2020-01-01", "2020-10-31"))%>%
  frameWidget(height =500)
```

What other visualizations can you make with these data?
