---
title: "Course Schedule"
output:
  html_document:
    toc: true
    toc_depth: 1
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
library(tidyverse)
source("functions.R")

yamls=append(
  yaml_dir(),
  yaml_dir(dir="pres"))

names=sapply(yamls,function(x) x$name)

#weeks=read_csv("schedule.csv")

sched=read_csv("schedule.csv")

```

Course content is designed to be flexible to accommodate student interest and abilities.  The order and timing of course topics may change as the semester progresses.  

**Readings**

We will refer to the following texts (all available online):

* R4DS [_R for Data Science_ by Garrett Grolemund and Hadley Wickham](http://r4ds.had.co.nz)
* GCR  [_Geocomputation with R_ by Robin Lovelace, Jakub Nowosad, and Jannes Muenchow (CRC Press, 2018)](https://geocompr.robinlovelace.net/)
* HG [Happy Git with R](https://happygitwithr.com/)


## Schedule

```{r table1,results='asis',echo=F}
cat("| Week | Date | Presentation | Tasks  |  Case Study | [DataCamp](https://www.datacamp.com/enterprise/geo511-fall-2020) | \n")
cat("|:-|:-|:-:|:---:|:---:|:---:|\n") 

not_na_exists<-function(x){
  if(is.na(x)) return("")
  if(!file.exists(x)) return("")
  else 
    return(x)
}

# link yamls to sched table by file name
yamlname=sapply(yamls,function(x) x$name)
sched$task_yaml=match(sched$task,yamlname)
sched$case_yaml=match(sched$case_study,yamlname)
sched$pres1_yaml=match(sched$pres1,yamlname)

for(i in sched$week){
  # task=ifelse(!is.na(sched$task[i]),paste0(sched$task[i],".Rmd"),NA)%>%
  #   not_na_exists()
  # case=ifelse(!is.na(sched$case_study[i]),
  #             paste0(sched$case_study[i],".Rmd"),NA)%>%
  #   not_na_exists()
  # pres1=ifelse(!is.na(sched$pres1[i]),
  #              paste0("pres/",sched$pres1[i],".html"),NA)%>%
  #   not_na_exists()
  # pres2=ifelse(!is.na(sched$pres1[i]),
  #              paste0("pres/",sched$pres2[i],".html"),NA)%>%
  #   not_na_exists()
  
  
  # ttask=ifelse(!is.na(task),yaml_extract(task),"")
  # tcase=ifelse(!is.na(case),yaml_extract(case),"")

  pres1=ifelse(
    grepl("http",sched$pres1[i]),
    sched$pres1[i],
    ifelse(
      is.null(yamls[[sched$pres1_yaml[i]]]$url),
      NA,
      yamls[[sched$pres1_yaml[i]]]$url))
 
  tpres_tb=ifelse(!is.na(pres1),
                   paste0( "[<i class='fas fa-desktop'>   </i>](",
                   pres1,"){target='_blank'}"),"")
 
  ttask_tb=paste0("[",yamls[[sched$task_yaml[i]]]$title,"](",
                  yamls[[sched$task_yaml[i]]]$url,")")
 
  tcase_tb=paste0("[",yamls[[sched$case_yaml[i]]]$title,"](",
                  yamls[[sched$case_yaml[i]]]$url,")")
  
  
  cat(paste(
    "| ",sched$week[i],
    "| ",sched$date[i],
    "| ",tpres_tb,
    "| ",ttask_tb,
    "| ",tcase_tb,
    "| ",ifelse(!is.na(sched$datacamp[i]),sched$datacamp[i],""),
    "|\n "
    ))
}
```

```{r table,results='asis',echo=F,eval=F}
# This chunk prints an index of all the task titles along with a preview of the sub-tasks 
cat("| Week | Date | Presentation | Tasks  |  Case Study | [Package Present](PackageIntro.html) |\n")
cat("|:--|:--|:-:|:-:|:-:|:---|:-:|\n") 
for(i in weeks$week){
  #print(i)
  # Find items for that week
yamlweeks=sapply(yamls,function(x) x$week)
yamltype=sapply(yamls,function(x) x$type)

wtasks=yamls[yamlweeks==i & yamltype=="Task"]
wcs=yamls[yamlweeks==i & yamltype=="Case Study"]
wpres=yamls[yamlweeks==i & yamltype=="Presentation"]

wtasks2=ifelse(length(wtasks)==1, 
               paste0("[",wtasks[[1]]$title,"](",wtasks[[1]]$url,")"), 
               " - ")

wcs2=ifelse(length(wcs)==1, 
               paste0("[",wcs[[1]]$title,"](",wcs[[1]]$url,")"), 
               " - ")
wpres2=ifelse(length(wpres)==1, 
               paste0( "[<i class='fas fa-desktop'>    </i>](",
                       wpres[[1]]$url,"){target='_blank'}"), 
               " - ")

  cat(paste(
    "| ",weeks$week[i],
    "| ",weeks$date[i],
#    wreadings,
    "| ",wpres2,
    "| ",wtasks2,
    "| ",wcs2,
    "| ",ifelse(is.na(weeks$package[i]),"",weeks$package[i]),
    "|\n "
    ))
}
```

## Student Leadership and Presentations

[Click here to edit the table below](https://docs.google.com/spreadsheets/d/1hH61DFHifYLdrbH3RQTIkR3zmo0ImpUg0tdHdBQF3Vw/edit?usp=sharing)


<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vSm_SdDqTPR5j7dXg_VeaUaKSxCVevvytsWQYKcT_kBCgOkKMsouHfafZ6wtbVjXaUSYxODReZgxAC3/pubhtml?gid=868447735&amp;single=true&amp;widget=true&amp;headers=false" style="height: 1000px; width: 100%;"></iframe>

