---
title: "Semester Wrap Up!"
type: Presentation
week: 13
---
```{r setup, message=FALSE, warning=F, echo=F, results='hide'}
knitr::opts_knit$set(root.dir = NULL)
knitr::opts_chunk$set(fig.width=800,fig.height=400)
library(tidyverse)
library(knitr)
```


## Course Schedule
<iframe
  src="http://wilsonlab.io/GEO511/Schedule.html"
  width="100%" height="800">
</iframe>
[source](http://wilsonlab.io/GEO511/Schedule.html)

## Case Study: Let's pick a winner!

<iframe
  src="https://wheelofnames.com/bw4-gdt"
  width="100%" height="500">
</iframe>
<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vSm_SdDqTPR5j7dXg_VeaUaKSxCVevvytsWQYKcT_kBCgOkKMsouHfafZ6wtbVjXaUSYxODReZgxAC3/pubhtml?gid=868447735&amp;single=true&amp;widget=true&amp;headers=false" style="height: 200px; width: 100%;"></iframe>


## Final Deliverables
Final deadline 1 week from last day of classes (December 18th).

<iframe
  src="http://wilsonlab.io/GEO511/Syllabus.html#semester-deliverables"
  width="100%" height="800">
</iframe>
[source](http://wilsonlab.io/GEO511/Syllabus.html#semester-deliverables)


# Final Projects

## Presentation Schedule
<iframe
  src="http://wilsonlab.io/GEO511/Student_projects.html#section"
  width="100%" height="800">
</iframe>
[source](http://wilsonlab.io/GEO511/Student_projects.html#section)

## Presentation Details

<iframe
  src="http://wilsonlab.io/GEO511/TK_15.html"
  width="100%" height="800">
</iframe>
[source](http://wilsonlab.io/GEO511/TK_15.html)


# Tips and Tricks

## Messages

If you have something like this in your website:
```
library(dplyr)
## 
## Attaching package: 'dplyr'
## The following objects are masked from 'package:stats':
## 
##     filter, lag
```

Remove/hide it with:

```
  {r message=F, warning=F, results="hide"}
``` 


## Code output

Don't simply print output!  Better to summarize everything as a table or figure. 

```{r}
lm(rnorm(10)~runif(10)) %>% 
  summary()
```

## Code output

Convert results to properly formatted output. 

```{r}
lm(rnorm(10)~runif(10)) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 2)
```



## Code folding?

If you have really long code chunks, you can add `code_folding: hide` to your YAML to hide all code by default.

```
output:
  html_document:
    code_folding: hide
```

## Chunks

Break up chunks with narrative to explain what the code is doing.  Chunks longer than ~20 lines should probably be broken up.

## Spelling

Don't forget to use spell check (Edit->Check Spelling)!  It's not automatic!

## And many more options to customize!
<iframe
  src="https://bookdown.org/yihui/rmarkdown-cookbook/html-scroll.html"
  width="100%" height="800">
</iframe>
[source](https://bookdown.org/yihui/rmarkdown-cookbook/html-scroll.html)


## Convert .Rmd to .R 

You can always convert a .Rmd file to a .R file (with or without comments) like this:

```
knitr::purl(inputfile, outputfile, documentation = 2)
```
