---
title: "Final Project Notes"
type: Presentation
week: 13
---
```{r setup, message=FALSE, warning=F, echo=F, results='hide'}
knitr::opts_knit$set(root.dir = NULL)
knitr::opts_chunk$set(fig.width=800,fig.height=400)
library(tidyverse)
library(knitr)
```

## Logistics
<iframe class='embed-responsive-item' src='../Schedule.html'  width=100% height=700px allowfullscreen></iframe>

## Resource Presentations

## Case Studies


# Project Suggestions

## Messages

If you have something like this in your website:
```
library(dplyr)
## 
## Attaching package: 'dplyr'
## The following objects are masked from 'package:stats':
## 
##     filter, lag
```

Remove/hide it with:

```
  {r message=F, warning=F, results="hide"}
``` 


## Code output

Don't simply print output!  Better to summarize everything as a table or figure. 

```{r}
lm(rnorm(10)~runif(10)) %>% 
  summary()
```

## Code output

Convert results to properly formatted output. 

```{r}
lm(rnorm(10)~runif(10)) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 2)
```



## Code folding?

If you have really long code chunks, you can add `code_folding: hide` to your YAML to hide all code by default.

```
output:
  html_document:
    code_folding: hide
```

## Chunks

Break up chunks with narrative to explain what the code is doing.  Chunks longer than ~20 lines should probably be broken up.

## Spelling

Don't forget to use spell check (Edit->Check Spelling)!  It's not automatic!


## Convert .Rmd to .R 

You can always convert a .Rmd file to a .R file (with or without comments) like this:

```
knitr::purl(inputfile, outputfile, documentation = 2)
```
