---
title: "Spatial data handling"
---

```{r, echo=F,message=F,warning=F}
library(methods)
library(rgdal)
library(rgeos)
library(sp)
library(sf)
library(tidyverse)
library(spData)
library(knitr)
library(ggplot2)
library(ggpubr)
```

## Course Schedule
<iframe
  src="http://wilsonlab.io/GEO511/Schedule.html"
  width="100%" height="800">
</iframe>
[source](http://wilsonlab.io/GEO511/Schedule.html)


## Resource Presentations

<iframe
  src="https://docs.google.com/spreadsheets/d/e/2PACX-1vSm_SdDqTPR5j7dXg_VeaUaKSxCVevvytsWQYKcT_kBCgOkKMsouHfafZ6wtbVjXaUSYxODReZgxAC3/pubhtml?gid=868447735&single=true"
  width="100%" height="800">
</iframe>
[source](https://docs.google.com/spreadsheets/d/e/2PACX-1vSm_SdDqTPR5j7dXg_VeaUaKSxCVevvytsWQYKcT_kBCgOkKMsouHfafZ6wtbVjXaUSYxODReZgxAC3/pubhtml?gid=868447735&single=true)

# Case Study Presentations

## Let's pick a winner!

<iframe
  src="https://wheelofnames.com/bw4-gdt"
  width="100%" height="500">
</iframe>
<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vSm_SdDqTPR5j7dXg_VeaUaKSxCVevvytsWQYKcT_kBCgOkKMsouHfafZ6wtbVjXaUSYxODReZgxAC3/pubhtml?gid=868447735&amp;single=true&amp;widget=true&amp;headers=false" style="height: 200px; width: 100%;"></iframe>

## Next Week's Case Study

<iframe
  src="http://wilsonlab.io/GEO511/CS_05.html"
  width="100%" height="800">
</iframe>
[source](http://wilsonlab.io/GEO511/CS_05.html)


---

# Spatial data handling in R

## Packages for Spatial Data

Packages:

* `sp`  First major spatial data package/format
* `sf` Spatial Features in the 'tidyverse'
* `rgdal` reading and writing spatial data
* `rgeos` Interface to open-source geometry engine (GEOS)
* `raster` raster (gridded) data
* and a few others...

---

## What is a Spatial Feature (sf)?

Typically a 'vector-based' represntation of an object in the real world:

```{r, echo=F}
px <- c(5, 7, 8, 9, 8, 7, 6) 
py <- c(7, 3, 4, 8, 9, 15, 14) 
plot(px, py, type="n", axes=F, xlab = '', ylab = '') 
polygon(px, py, col = "khaki1")
points(c(6, 9, 8, 8.5), c(9, 14, 8, 9), pch=20, col = "peachpuff4", lwd = 3)
lines(c(5, 6, 7, 8), c(5, 6,10, 11), col = "steelblue1",  lwd = 3)
lines(c(8, 9), c(14, 12), col = "dark green", lwd = 3)
```

## Geometry

Features have a _geometry_ describing _where_ on Earth the
feature is located, and they have attributes, which describe other
properties. 

A Tree: 

* delineation of its crown
* its stem
* point indicating its centre

Attributes:

* species 
* height
* diameter
* date


---

## Spatial Features

What information do we need to store in order to define points, lines, polygons in geographic space?

* lat ($y$)
* lon ($x$)
* elevation/altitude ($z$)
* measure ($m$, rarely used) associated with the point, rather than with the feature as a whole (e.g. coordinate measurement error)
 - projection
 - what type (point/line/poly)
 - if polygon, is it a hole or not
 - attribute data
 - ... ?


# `sp` package

---

##  [`sp`](https://cran.r-project.org/package=sp) Package

* First to provide classes and methods to create spatial _points_, _lines_, _polygons_, and _grids_.
* ~350 spatial analysis packages use `sp` data types
* Important to know how to convert **sp** to and from **sf** objects
* Recommend using `sf` methods instead and convert only when necessary.

# `sf` Package


## Simple feature geometry types

The following seven simple feature types are the most common, and are for instance the only ones used for [GeoJSON](https://tools.ietf.org/html/rfc7946):

| type | description |
| ---- | ----------- |
| `POINT` | a single point |
| `LINESTRING` | sequence of points connected by lines |
| `POLYGON` | sequence of points form a closed ring |
| `MULTIPOINT` | set of points |
| `MULTILINESTRING` | set of linestrings |
| `MULTIPOLYGON` | set of polygons |
| `GEOMETRYCOLLECTION` | set of geometries |


## Uncommon Geometry Types 
10 more geometries 10 are rare:

* `CIRCULARSTRING`
* `COMPOUNDCURVE` 
* `CURVEPOLYGON` 
* `MULTICURVE` 
* `MULTISURFACE` 
* `CURVE` 
* `SURFACE` 
* `POLYHEDRALSURFACE` 
* `TIN` 
* `TRIANGLE`

---

## Coordinate reference system

Coordinates only have meaning on the Earth's surface when their coordinate
reference system (CRS) is known. 

Overview of [projection systems](https://www.e-education.psu.edu/natureofgeoinfo/c2_p28.html)

---

## Reproject to ["Goode Homolosine" projection](https://proj.org/operations/projections/goode.html)
```{r, warning=F}
data(world) # from spData package
st_crs(world)

world_homolosine=world %>% 
  st_transform(crs = "+proj=goode")

p1 <- ggplot(world,aes(fill=gdpPercap))+
  geom_sf()
p2 <- ggplot(world_homolosine,aes(fill=gdpPercap))+
  geom_sf()
```

## 
```{r}
ggarrange(p1,p2,common.legend = T)
```

## Defining a projection

* EPSG code
   * `EPSG:4326`
* Proj4 String
   * `+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs` 
* WKT (well known text) string
   * `GEOGCS["WGS 84",
    DATUM["WGS_1984",
        SPHEROID["WGS 84",6378137,298.257223563,
            AUTHORITY["EPSG","7030"]],
        AUTHORITY["EPSG","6326"]],
    PRIMEM["Greenwich",0,
        AUTHORITY["EPSG","8901"]],
    UNIT["degree",0.01745329251994328,
        AUTHORITY["EPSG","9122"]],
    AUTHORITY["EPSG","4326"]]`

[spatialreference.org](https://spatialreference.org/ref/epsg/wgs-84/prettywkt/)
[proj.org](https://proj.org/operations/projections/index.html)

---

## How simple features in R are organized

Like [PostGIS](http://postgis.net/), all functions in `sf` prefixed by `st_` (_spatial and temporal_).
Easily findable by command-line completion.  

---

* Objects stored in `data.frame` objects (or `tbl_df`)
* Feature geometries in a single `data.frame` column (as a list). 
* Each list element holding the simple feature geometry of that feature (e.g. point, polygon)

---

```{r}
glimpse(world)
```

---

## Components

* `sf`, the table (`data.frame`) with feature attributes and feature geometries, which contains
* `sfc`, the list-column with the geometries for each feature (record), which is composed of
* `sfg`, the feature geometry of an individual simple feature.


---

If you work with PostGis or GeoJSON you may have come across the [WKT (well-known text)](https://en.wikipedia.org/wiki/Well-known_text) format, for example like these: 

    `POINT (30 10)
    LINESTRING (30 10, 10 30, 40 40)
    POLYGON ((30 10, 40 40, 20 40, 10 20, 30 10))`

`sf` implements this standard natively in R. Data are structured and conceptualized very differently from the `sp` approach.

---

## sf Highlights

* provides **fast** I/O, particularly relevant for large files 
* directly reads from and writes to spatial **databases** such as PostGIS
* compatibile with the *tidyverse*
* `ggplot` compatible with `geom_sf()`

---

`sp` and `sf` are _not_ only formats for spatial objects. Other spatial packages may use their own class definitions for spatial data (for example `spatstat`). Usually you can find functions that convert `sp` and increasingly `sf` objects to and from these formats.

---

```{r}
world_sp = as(world, "Spatial")
world_sf = st_as_sf(world_sp)
str(world_sp)
```

---

```{r}
str(world)
```

---



- The structures in the **sp** packages are more complicated - `str(world_sf)` vs `str(world_sp)`
  
- Moreover, many of the **sp** functions are not "pipeable" (it's hard to combine them with the **tidyverse**)

```{r, eval = F}
world_sp %>% 
  filter(name_long == "United Kingdom")
```

`Error in UseMethod("filter_") : 
  no applicable method for 'filter_' applied to an object of class "c('SpatialPolygonsDataFrame', 'SpatialPolygons', 'Spatial', 'SpatialPolygonsNULL', 'SpatialVector')"`

---
```{r, eval = F}
world %>% 
  filter(name_long == "United Kingdom")%>%
  ggplot()+geom_sf()
```

---

## Reading and writing spatial data

```{r}
vector_filepath = system.file("shapes/world.gpkg", package = "spData")
vector_filepath
world = st_read(vector_filepath)
```

Counterpart to `st_read()` is the `st_write` function, e.g. `st_write(world, 'data/new_world.gpkg')`. A full list of supported formats could be found using `sf::st_drivers()`.

---

## Structure of the sf objects 

```{r, eval = FALSE}
world
```

```{r, echo = FALSE}
print(world, n=3)
```

```{r}
class(world)
```

---

## Structure of the sf objects 

```{r, eval=FALSE}
world$name_long
```

```{r, echo=FALSE}
world$name_long[1:3]
```

```{r, eval=FALSE}
world$geom
```

```{r, echo=FALSE}
print(world$geom, n = 3)
```


## Non-spatial operations on the sf objects

```{r, warning=FALSE}
world %>% 
  left_join(worldbank_df, by = "iso_a2") %>%
  select(name_long, pop, pop_growth, area_km2) %>%
  mutate(pop_density = pop/area_km2) %>%
  rename(population = pop) %>% 
  arrange(desc(pop_density))%>%
  kable()
```

---

## Non-spatial operations

```{r}
world_cont = world %>% 
        group_by(continent) %>% 
        summarize(pop_sum = sum(pop, na.rm = TRUE))
```

```{r, echo=FALSE}
print(world_cont, n = 1)

ggplot(world_cont,aes(fill=pop_sum))+
  geom_sf()
```

---

The `st_set_geometry` function can be used to remove the geometry column:

```{r}
world_df =st_set_geometry(world_cont, NULL)
class(world_df)
```

---

## Spatial operations

It's a big topic which includes:

- Spatial subsetting
- Spatial joining/aggregation
- Topological relations
- Distances
- Spatial geometry modification
- Raster operations (map algebra)

See [Chapter 4](http://robinlovelace.net/geocompr/spatial-data-operations.html#spatial-operations-on-raster-data) of *Geocomputation with R* 

## CRS

Transform (warp) to a different projection:
```{r}
na_2163 = world %>%
  filter(continent == "North America") %>% 
  st_transform(2163) #US National Atlas Equal Area
st_crs(na_2163)
```

---

## Compare projections

```{r}
na = world %>%
  filter(continent == "North America") 
p1=ggplot(na)+geom_sf()
p2=ggplot(na_2163)+geom_sf()
ggarrange(p1,p2)
```

---

##  The `world` dataset

```{r}
plot(na_2163)
```

---

## Spatial operations

```{r, warning = FALSE, message = FALSE, fig.height = 4}
canada = na_2163 %>% 
  filter(name_long=="Canada")

canada_buffer=canada%>%
  st_buffer(500000) %>%
  filter()

ggplot()+
  geom_sf(data=canada)+
  geom_sf(data=canada_buffer,col="red",fill=NA)
```


## Basic maps

- Basic maps of `sf` objects can be quickly created using the `plot()` function:

```{r}
plot(world[0])
```

## Basic maps

```{r}
plot(world["pop"])
```

---

## leaflet
```{r}
library(leaflet)
library(widgetframe)
```

```{r, eval=F}
leaflet(world) %>%
        addTiles() %>%
        addPolygons(color = "#444444", weight = 1, fillOpacity = 0.5,
                    fillColor = ~colorQuantile("YlOrRd", lifeExp)(lifeExp),
                    popup = paste(round(world$lifeExp, 2)))
```

```{r, echo=T, message=T,results='asis'}
l = leaflet(world) %>%
  addTiles() %>%
  addPolygons(color = "#444444", 
              weight = 1, fillOpacity = 0.5, 
              fillColor = ~colorQuantile("YlOrRd", lifeExp)(lifeExp),
              popup = paste(round(world$lifeExp, 2)))
saveWidget(l, file = "leaflet.html",selfcontained = T)
```
<iframe src="leaflet.html" width=100% height=400px></iframe>
---

## Raster data in the tidyverse

Raster data is not yet closely connected to the **tidyverse**, however:

- Some functions from the **raster** package works well in `pipes`
- You can convert vector data to the `Spatial*` form using `as(my_vector, "Spatial")`before working on raster-vector interactions
- There are some initial efforts to bring raster data closer to the **tidyverse**, such as [tabularaster](https://github.com/hypertidy/tabularaster), [sfraster](https://github.com/mdsumner/sfraster), or [fasterize](https://github.com/ecohealthalliance/fasterize)
- The development of the [stars](https://github.com/r-spatial/stars), package focused on multidimensional, large datasets should start soon. It will allow pipe-based workflows

---

## Sources

- Slides adapted from:
    - "Robin Lovelace and Jakub Nowosad" draft book [_Geocomputation with R_ (to be published in 2018)](http://robinlovelace.net/geocompr/). Source code at https://github.com/robinlovelace/geocompr.
    - [Claudia Engel's spatial analysis workshop](https://github.com/cengel/rspatial/blob/master/2_spDataTypes.Rmd)

