---
title: "Final Project Notes"
type: Presentation
week: 13
---
```{r setup, echo=F, results='hide'}
knitr::opts_knit$set(root.dir = NULL)
knitr::opts_chunk$set(fig.width=800,fig.height=400)
library(tidyverse)
library(knitr)
```

## Logistics
<iframe class='embed-responsive-item' src='../Schedule.html'  width=100% height=700px allowfullscreen></iframe>

## Resource Presentations

## Case Studies


# Project Suggestions

## Messages


```
library(dplyr)
## 
## Attaching package: 'dplyr'
## The following objects are masked from 'package:stats':
## 
##     filter, lag
```

Use:
```
  {r message=F, warning=F, results="hide"}
``` 

to hide output.


## Code output

Best to summarize everything as a table or figure - don't simply print output:

```{r}
lm(rnorm(10)~runif(10)) %>% 
  summary()
```

## Code output

```{r}
lm(rnorm(10)~runif(10)) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 2)
```

Convert results to properly formatted output. 

## Code folding?

Add `code_folding: hide` to your YAML to hide all code by default.

```
output:
  html_document:
    code_folding: hide
```

## Chunks

Break up chunks with narrative to explain what the code is doing.

## Spelling

Use spell check (Edit->Check Spelling!


## Convert .Rmd to .R 

You can always convert a .Rmd file to a .R file (with or without comments) like this:

```
knitr::purl(inputfile, outputfile, documentation = 2)
```
