---
title: Satellite Remote Sensing
subtitle: Analyze Satellite Data
week: 10
type: Case Study
reading:
tasks:
   - Write a .Rmd script to perform the following tasks
   - 
   - 
---

```{r, echo=FALSE, message=FALSE, results='hide', purl=FALSE}
source("knitr_header.R")
```


[<i class="fa fa-file-code-o fa-3x" aria-hidden="true"></i> The R Script associated with this page is available here](`r output`).  If you like, you can download this file and open it (or copy-paste into a new script) with RStudio so you can follow along.  


### Libraries

```{r,results='hide',message=FALSE, warning=F}
library(raster)
library(rasterVis)
library(rgdal)
library(ggmap)
library(tidyverse)
library(knitr)

# New Packages
library(ncdf4) # to import data from netcdf format
```


## Identify (and create) download folders

Today we'll work with:

* Land Surface Temperature (`lst`): MOD11A2
* Land Cover (`lc`): MCD12Q1

## Land Use Land Cover

```{r, eval=F}
# Create afolder to hold the downloaded data
dir.create("data",showWarnings = F) #create a folder to hold the data

lulc_url="https://github.com/adammwilson/DataScienceData/blob/master/inst/extdata/appeears/MCD12Q1.051_aid0001.nc?raw=true"
lst_url="https://github.com/adammwilson/DataScienceData/blob/master/inst/extdata/appeears/MOD11A2.006_aid0001.nc?raw=true"

# download them
download.file(lulc_url,destfile="data/MCD12Q1.051_aid0001.nc")
download.file(lst_url,destfile="data/MOD11A2.006_aid0001.nc")
```

You should also edit your .gitignore file (in your tasks repository folder) to include `*data*` on one line. This will prevent git from adding these files.  


## Load data into R
```{r}
lulc=stack("data/MCD12Q1.051_aid0001.nc",varname="Land_Cover_Type_1")
lst=stack("data/MOD11A2.006_aid0001.nc",varname="LST_Day_1km")
```

You will probably see some errors about 

```
>>>> WARNING <<<  attribute longitude_of_projection_origin is an 8-byte value, but R"
[1] "does not support this data type. I am returning a double precision"
[1] "floating point, but you must be aware that this could lose precision!"
```

You can safely ignore this.  


## Explore LULC data
```{r, warning=F, message=FALSE,results='hide'}
plot(lulc)
```
 

We'll just pick one year to work with to keep this simple:
```{r, warning=F}
lulc=lulc[[13]]
plot(lulc)
```

### Process landcover data

Assign land cover clases from [MODIS website](https://lpdaac.usgs.gov/dataset_discovery/modis/modis_products_table/mcd12q1)

```{r}
  Land_Cover_Type_1 = c(
    Water = 0, 
    `Evergreen Needleleaf forest` = 1, 
    `Evergreen Broadleaf forest` = 2,
    `Deciduous Needleleaf forest` = 3, 
    `Deciduous Broadleaf forest` = 4,
    `Mixed forest` = 5, 
    `Closed shrublands` = 6,
    `Open shrublands` = 7,
    `Woody savannas` = 8, 
    Savannas = 9,
    Grasslands = 10,
    `Permanent wetlands` = 11, 
    Croplands = 12,
    `Urban & built-up` = 13,
    `Cropland/Natural vegetation mosaic` = 14, 
    `Snow & ice` = 15,
    `Barren/Sparsely vegetated` = 16, 
    Unclassified = 254,
    NoDataFill = 255)

lcd=data.frame(
  ID=Land_Cover_Type_1,
  landcover=names(Land_Cover_Type_1),
  col=c("#000080","#008000","#00FF00", "#99CC00","#99FF99", "#339966", "#993366", "#FFCC99", "#CCFFCC", "#FFCC00", "#FF9900", "#006699", "#FFFF00", "#FF0000", "#999966", "#FFFFFF", "#808080", "#000000", "#000000"),
  stringsAsFactors = F)
# colors from https://lpdaac.usgs.gov/about/news_archive/modisterra_land_cover_types_yearly_l3_global_005deg_cmg_mod12c1
kable(head(lcd))
```        

Convert LULC raster into a 'factor' (categorical) raster.  This requires building the Raster Attribute Table (RAT).  Unfortunately, this is a bit of manual process as follows.
```{r}
# convert to raster (easy)
lulc=as.factor(lulc)

# update the RAT with a left join
levels(lulc)=left_join(levels(lulc)[[1]],lcd)

```

```{r, fig.height=12, warning=F, purl=F}
# plot it
gplot(lulc)+
  geom_raster(aes(fill=as.factor(value)))+
  scale_fill_manual(values=levels(lulc)[[1]]$col,
                    labels=levels(lulc)[[1]]$landcover,
                    name="Landcover Type")+
  coord_equal()+
  theme(legend.position = "right")+
  guides(fill=guide_legend(ncol=1,byrow=TRUE))
```

# Land Surface Temperature
```{r, warning=F, message=FALSE,results='hide', purl=F}
plot(lst[[1:12]])
```

## Convert LST to Degrees C 
You can convert LST from Degrees Kelvin (K) to Celcius (C) with `offs()`.

```{r}
offs(lst)=-273.15
plot(lst[[1:10]])
```


<div class="well">

# MODLAND Quality control

See a detailed explaination [here](https://lpdaac.usgs.gov/sites/default/files/public/modis/docs/MODIS_LP_QA_Tutorial-1b.pdf).  Some code below from [Steven Mosher's blog](https://stevemosher.wordpress.com/2012/12/05/modis-qc-bits/).

Expand this to learn more about MODIS quality control.

<button data-toggle="collapse" class="btn btn-primary btn-sm round" data-target="#demo1">More Info</button>
<div id="demo1" class="collapse">
## MOD11A2 (Land Surface Temperature) Quality Control

[MOD11A2 QC Layer table](https://lpdaac.usgs.gov/dataset_discovery/modis/modis_products_table/mod11a2)

```{r, warning=F, message=FALSE,results='hide', purl=F}
lstqc=stack("data/MOD11A2.006_aid0001.nc",varname="QC_Day")
plot(lstqc[[1:2]])
```

### LST QC data

QC data are encoded in 8-bit 'words' to compress information.

```{r, purl=F}
values(lstqc[[1:2]])%>%table()
```

```{r, purl=F}
intToBits(65)
intToBits(65)[1:8]

as.integer(intToBits(65)[1:8])
```
#### MODIS QC data are _Big Endian_

Format          Digits              value     sum
----            ----                ----      ----
Little Endian   1 0 0 0 0 0 1 0     65        2^0 + 2^6
Big Endian      0 1 0 0 0 0 0 1     65        2^6 + 2^0


Reverse the digits with `rev()` and compare with QC table above.

```{r, purl=F}
rev(as.integer(intToBits(65)[1:8]))
```
QC for value `65`:

* LST produced, other quality, recommend examination of more detailed QA
* good data quality of L1B in 7 TIR bands
* average emissivity error <= 0.01
* Average LST error <= 2K

### Filter the the lst data using the QC data

```{r, purl=F}
## set up data frame to hold all combinations
QC_Data <- data.frame(Integer_Value = 0:255,
Bit7 = NA, Bit6 = NA, Bit5 = NA, Bit4 = NA,
Bit3 = NA, Bit2 = NA, Bit1 = NA, Bit0 = NA,
QA_word1 = NA, QA_word2 = NA, QA_word3 = NA,
QA_word4 = NA)

## 
for(i in QC_Data$Integer_Value){
AsInt <- as.integer(intToBits(i)[1:8])
QC_Data[i+1,2:9]<- AsInt[8:1]
}

QC_Data$QA_word1[QC_Data$Bit1 == 0 & QC_Data$Bit0==0] <- "LST GOOD"
QC_Data$QA_word1[QC_Data$Bit1 == 0 & QC_Data$Bit0==1] <- "LST Produced,Other Quality"
QC_Data$QA_word1[QC_Data$Bit1 == 1 & QC_Data$Bit0==0] <- "No Pixel,clouds"
QC_Data$QA_word1[QC_Data$Bit1 == 1 & QC_Data$Bit0==1] <- "No Pixel, Other QA"

QC_Data$QA_word2[QC_Data$Bit3 == 0 & QC_Data$Bit2==0] <- "Good Data"
QC_Data$QA_word2[QC_Data$Bit3 == 0 & QC_Data$Bit2==1] <- "Other Quality"
QC_Data$QA_word2[QC_Data$Bit3 == 1 & QC_Data$Bit2==0] <- "TBD"
QC_Data$QA_word2[QC_Data$Bit3 == 1 & QC_Data$Bit2==1] <- "TBD"

QC_Data$QA_word3[QC_Data$Bit5 == 0 & QC_Data$Bit4==0] <- "Emiss Error <= .01"
QC_Data$QA_word3[QC_Data$Bit5 == 0 & QC_Data$Bit4==1] <- "Emiss Err >.01 <=.02"
QC_Data$QA_word3[QC_Data$Bit5 == 1 & QC_Data$Bit4==0] <- "Emiss Err >.02 <=.04"
QC_Data$QA_word3[QC_Data$Bit5 == 1 & QC_Data$Bit4==1] <- "Emiss Err > .04"

QC_Data$QA_word4[QC_Data$Bit7 == 0 & QC_Data$Bit6==0] <- "LST Err <= 1"
QC_Data$QA_word4[QC_Data$Bit7 == 0 & QC_Data$Bit6==1] <- "LST Err > 2 LST Err <= 3"
QC_Data$QA_word4[QC_Data$Bit7 == 1 & QC_Data$Bit6==0] <- "LST Err > 1 LST Err <= 2"
QC_Data$QA_word4[QC_Data$Bit7 == 1 & QC_Data$Bit6==1] <- "LST Err > 4"
kable(head(QC_Data))
```

### Select which QC Levels to keep
```{r, purl=F}
keep=QC_Data[QC_Data$Bit1 == 0,]
keepvals=unique(keep$Integer_Value)
keepvals
```

### How many observations will be dropped?

```{r,warning=F, purl=F}
qcvals=table(values(lstqc))  # this takes a minute or two


QC_Data%>%
  dplyr::select(everything(),-contains("Bit"))%>%
  mutate(Var1=as.character(Integer_Value),
         keep=Integer_Value%in%keepvals)%>%
  inner_join(data.frame(qcvals))%>%
  kable()
```

Do you want to update the values you are keeping?

### Filter the LST Data keeping only `keepvals`

These steps take a couple minutes.  

Make logical flag to use for mask
```{r, purl=F}
lstkeep=calc(lstqc,function(x) x%in%keepvals)
```

Plot the mask
```{r,fig.height=12, purl=F}
gplot(lstkeep[[4:8]])+
  geom_raster(aes(fill=as.factor(value)))+
  facet_grid(variable~.)+
  scale_fill_manual(values=c("blue","red"),name="Keep")+
  coord_equal()+
  theme(legend.position = "bottom")
```


Mask the lst data using the QC data and overwrite the original data.
```{r, purl=F}
lst=mask(lst,mask=lstkeep,maskval=0)
```

</div>
</div>



## Add Dates to Z (time) dimension

The default layer names of the LST file include the date as follows:

```{r, purl=F}
names(lst)[1:5]
```

Convert those values to a proper R Date format by dropping the "X" and using `as.Date()`.
```{r}
tdates=names(lst)%>%
  sub(pattern="X",replacement="")%>%
  as.Date("%Y.%m.%d")

names(lst)=1:nlayers(lst)
lst=setZ(lst,tdates)
```


## Extract timeseries for a point

Define a point and transform it to the same projection as the LST data.

<div class="well">
## Hints
Use `stackApply()` to generate and plot monthly median lst values.

Hints:

<button data-toggle="collapse" class="btn btn-primary btn-sm round" data-target="#demo2">Hints</button>
<div id="demo2" class="collapse">

1. Use `lw=SpatialPoints(data.frame(x= -78.791547,y=43.007211))` to define a new Spatial Point at that location.
2. Set the projection of your point with `projection()` to `"+proj=longlat"`.
3. Transform the point to the projection of the raster using `spTransform()`.
4. Extract the LST data for that location with: `extract(lst,lw,buffer=1000,fun=mean,na.rm=T)`
5. Plot it with `ggplot()` including points for the raw data and a smooth version as a line.  You will probably want to adjust both `span` and `n` in `geom_smooth`.

</div>
</div>

Your graph should look like this:
```{r, warning=F, purl=F}
lw=SpatialPoints(
  data.frame(
    x= -78.791547,
    y=43.007211))
projection(lw)="+proj=longlat"

lw=spTransform(lw,projection(lst))

lwt=data.frame(date=getZ(lst),
                 lst=t(raster::extract(
                   lst,lw,
                   buffer=1000,
                   fun=mean,na.rm=T)))

ggplot(lwt,aes(x=date,y=lst))+
  geom_point()+
  geom_smooth(span=.05,n=1000)
```

See the `library(rts)` for more timeseries related functions.


# Summarize weekly data to Monthly Climatologies

Now we will use a function called `stackApply()` with a seasonal index to calculate monthly mean land surface temperature.

<div class="well">
## Hints
Use `stackApply()` to generate and plot monthly median lst values.

Hints:

1. First make a variable called `tmonth` by converting the dates to months using `as.numeric(format(getZ(lst2),"%m"))`
2. Use `stackApply()` to summarize the mean value per month (using the `tmonth` variable you just created) and save the results as `lst_month`.
3. Set the names of the layers to months with `names(lst_month)=month.name`
4. Plot the map for each month with `gplot()` in the RasterVis Package.
5. Calculate the monthly mean for the entire image with `cellStats(lst_month,mean)`

<button data-toggle="collapse" class="btn btn-primary btn-sm round" data-target="#demo2">Hints</button>
<div id="demo2" class="collapse">


The solution will look like this:
```{r, purl=F}
tmonth=as.numeric(format(getZ(lst),"%m"))

lst_month=stackApply(lst,indices = tmonth,mean,na.rm=T)
names(lst_month)=month.name

gplot(lst_month)+geom_raster(aes(fill=value))+
  facet_wrap(~variable)+
  scale_fill_gradientn(colours=c("blue",mid="grey","red"))+
  coord_equal()
```

And the table should be as follows:
```{r, purl=F}
monthly=cellStats(lst_month,mean) %>% round(2)
kable(monthly)
```

</div>
</div>


## Combine Land Cover and LST data


<div class="well">

## Hints
1. Resample `lc` to `lst` grid
2. Summarize mean monthly temperatures by Landcover
3. Extract values from `lst` and `lc` rasters.  
4. Melt table and add LandCover Name

<button data-toggle="collapse" class="btn btn-primary btn-sm round" data-target="#demo2">Hints</button>
<div id="demo2" class="collapse">


```{r}
lulc2=resample(lulc,
             lst,
             method="ngb")

lcds1=cbind.data.frame(
  values(lst_month),
  ID=values(lulc2[[1]]))%>%
  na.omit()

lcds2=lcds1%>%
  gather(key="month", value = "value", -ID)%>%
  mutate(ID=as.numeric(ID),
         month=factor(month,levels=month.name,ordered=T))%>%
  left_join(lcd)
head(lcds2)
```

#### Explore LST distributions by landcover

```{r,fig.height=12}
lcds2 %>% 
  filter(landcover%in%c("Urban & built-up","Deciduous Broadleaf forest")) %>% 
  ggplot(aes(y=value,x=month))+
  facet_wrap(~landcover)+
  geom_point(alpha=.5,position="jitter")+
  geom_smooth()+
  geom_violin(alpha=.5,col="red",scale = "width",position="dodge")+
  theme(axis.text.x=element_text(angle=90, hjust=1))+
  ylab("Monthly Mean Land Surface Temperature (C)")+
  xlab("Month")+
  ggtitle("Comparison of Land Surface Temperature in Urban and Forest areas in Buffalo, NY")
```


### Use Zonal Statistics to calculate summaries
```{r}
lct.mean=raster::zonal(lst_month,
               lulc2,
               'mean',na.rm=T)%>%
  data.frame()

lct.sd=zonal(lst_month,
             lulc2,
             'sd',na.rm=T)%>%
  data.frame()

lct.count=zonal(lst_month,
                lulc2,
                'count',na.rm=T)%>%
  data.frame()

lct.summary=rbind(data.frame(lct.mean,var="mean"),
                  data.frame(lct.sd,var="sd"),
                  data.frame(lct.count,var="count"))
```

#### Summarize seasonal values
```{r}
lctl=gather(lct.summary, key="month", value="value", -var, -zone)

lctl$month=factor(lctl$month,
                   labels=month.name,
                   ordered=T)
lctl$zone=names(Land_Cover_Type_1)[lctl$zone+1]
lctl=spread(lctl,var,value="value")
head(lctl)%>%kable()
```

## Build summary table
```{r}
filter(lctl,count>=100)%>%
  mutate(txt=paste0(round(mean,2),
                    " (±",round(sd,2),")"))%>%
  dplyr::select(zone,count,txt,month)%>%
  spread(month, txt)%>%
  kable()
```

<div class="well">
## Your turn
Calculate the  maximum observed seasonal average lst in each land cover type.  
Hints:

1. First use `zonal()` of `lst_seas` and `lulc2` to calculate the `max()` with `na.rm=T`
2. convert the output to a `data.frame()`
3. use `arrange()` to sort by `desc(max)`
4. use `kable()` if desired to make it print nicely.

<button data-toggle="collapse" class="btn btn-primary btn-sm round" data-target="#demo3">Show Solution</button>
<div id="demo3" class="collapse">
```{r, purl=F}
zonal(max(lst_seas),lulc2,'max',na.rm=T)%>%
  data.frame()%>%
  arrange(desc(max))%>%
  kable()
```

</div>
</div>



Things to think about:

* What tests would you use to identify differences?
* Do you need to worry about unequal sample sizes?

